---
output:
  html_document:
    df_print: paged
---



## Intervention by reducing $c$ and increasing $f$ 
<font size="3">
In the intervention model you assumed **`r sprintf("%.2f", ((F._Value)*100))`** % of population reduce their social contacts by **`r sprintf("%.2f", (1-C.)*100 )`**%. With such intervention the individual social contacts will reduce  to **`r sprintf("%.2f", (1-Opt()$par[4])*100-((1-C.)*100))`** % for **`r sprintf("%.2f", (Opt()$par[5])*100+((F._Value)*100))`** % of population. The max number of confirmed cases (I.i) will reach to **`r sprintf("%.0f", ImaxInterv)`** on **`r as.character(DateMaxInterv)`**.`r ifelse(is.na(INextmaxInterv),print("This max value is at the upper bound of the data, and therefore may not be the actual max, increase forecast step to reach the peak value."),print(""))` This intervention gives an effective reproduction number ($R_t$) about **`r sprintf("%.0f", (((1-(1-C.^2)*F._Value)*R_zero)/R_zero)*100)`**% of the basic reproduction number ($R_0$).


<font>


```{r,echo=F, results='asis'}

BETA<-Opt()$par[1]*(1-(1-C.^2)*F._Value)
SIGMA<-Opt()$par[3]
GAMA<-Opt()$par[2]
R0<-Opt()$par[1]/Opt()$par[2]
RT<-(1-(1-C.^2)*F._Value)*R_zero
R0R<-sprintf("%.0f", 100-((((1-(1-C.^2)*F._Value)*R_zero)/R_zero)*100))
C.value<-(1-Opt()$par[4])*100-((1-C.)*100)
F.value<- (Opt()$par[5])*100+((F._Value)*100)
Population<-N.
Ib<-sprintf("%.0f", Imax)
Ii<-sprintf("%.0f", ImaxInterv)
Dateb<- as.character(DateMax)
Datei<-as.character(DateMaxInterv)
  

 a <- structure(list( c("$\\beta$", "$\\sigma$", "$\\gamma$", "$R_0$", "$R_t$","$R_r$","$c$","$f$","$N$","$I_b$","$Date_{I_b}$","$I_i$","$Date_{I_i}$"
), Definition = c(
                  "the effective contact rate",
                  "The infection rate of an exposed person",
                  "The recovery rate of an infected person",
                  "The basic reproduction number",
                  "The effective reproduction number",
                  "Reduction of basic reproduction number",
                  "The social contacts",
                  "Population involved in social distancing intervention", 
                  "Population",
                  "The basic number of confirmed cases",
                  "$I_b$ peak date",
                  "Confirmed cases with intervention",
                  "$I_i$ peak date"), Unit = c("$day^{-1}$", 
"$day^{-1}$", "$day^{-1}$", "dimensionless", "dimensionless","%","%","%","People","People","Date","People","Date"),Value = c(BETA, SIGMA, GAMA, R0,RT,R0R,C.value,F.value,Population,Ib,Dateb,Ii,Datei)), .Names = c("Parameter","Definition", "Unit", "Estimated value"), row.names = c(NA, 13L), class = "data.frame")
 

```

```{r,echo=F, eval=T,out.width = "30%"}
 kable(a, booktabs = TRUE, longtable = TRUE,caption =  "Estimated values of epidemiological parameters in SEIR model") %>% 
  kable_styling(full_width = F)  
   
```


